An investigation into the contentious phylogenetic results of the *Mixta* species

### 2021-04-20 Manuscript/Manuscript_Paper.Rmd
Update Manuscript_Paper.Rmd

---
title: "An investigation into the contentious phylogenetic results of the *Mixta* species"
output:
  word_document:
    reference_docx: Style_Manuscript.docx
    fig_caption: true
    # toc: true
    # toc_depth: 5
bibliography: References.bib
csl: IJSEM.csl
---
```{r Packages, echo = FALSE, message = FALSE, warning=FALSE}
library("knitr")
library("tidyverse")
library("ggplot2")
library("RColorBrewer")
library("extrafont")
theme_set(theme_classic())
```

@@

The closest relatives to both of the *Mixta* species were extracted from RM and BAM distance matrices. When the RM was used. *P. septica* was the closest relative to 66.3 and 66.5% of the *M. calida* and *M. gaviniae* sequences, respectively (Figure 1). These percentages increase by 12.6 and 14.4% for the *M. calida* and *M. gaviniae* sequences, respectively, when the BAM was used (Figure 1). Instead of *P. septica* when the RM is used, more genes are more closely related to *P. agglomerans*, *E. tasmaniensis*, *E. amylovora*, *T. saanichensis*, and *E. cloacae* (Figure 1). Additionally, there was more disparity between *M. calida* and *M. gaviniae* when the BAM was used (Figure 1).

@@

## Closest Relative from Genetic Distance
Genetic distance gives a numerical estimate of evolutionary change between species, and the taxa with which a species of interest has the shortest distance is regarded as the closest relative. Since the phylogenetic trees were calculated using the RM, the distance matrices exported from the trees were kept for future analyses. *P. septica* was consistently the closest relative for most of the *Mixta* sequences (Figure 2). However, the number of genes most closely related *P. septica* decreased from 66.3 and 66.5% (*M. calida* and *M. gaviniae*, respectively) with the NTS by 23.7% with the amino acid sequences (AAS; Figure 2). Instead, the *Mixta* AAS were most closely related to the other taxa within the dataset except for *P. syringae* (Figure 2). Interestingly, the number of genes most closely related to *E. cloacae* increased between the NTS and the AAS from 2.9% by 3.1% (Figure 2).

@@

@@

```{r Figure1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 1. The percent of genes (n = 799) requiring each model when using the recommended model for phylogenetic tree analysis (green) and when using the best available model for genetic distance analysis (orange). Models are for nucleotide sequences.", fig.width=6.5}
models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_BestvAvailableModels.csv", stringsAsFactors = FALSE)

ggplot(data = models, aes(x = Model, y = Percent, fill = Bst_Avlb)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Dark2",
                    labels = c("Phylogenetic", "Genetic Distance")) +
  labs(y = "Percent of Genes \n Requiring Each Model",
       fill = "Analysis") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),                                # Rotates x axis labels
        legend.position = c(0.3, 0.95),                                                   # Moves legend inside the plot and in the top-left corner
        legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates
        legend.text.align = 0,                                                            # Aligns legend text to the left
        legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend
        text = element_text(size = 12,  family = "Times New Roman"))
```

>>>>

---
title: "An investigation into the contentious phylogenetic results of the *Mixta* species"
output:
  word_document:
    reference_docx: Style_Manuscript.docx
    fig_caption: true
    # toc: true
    # toc_depth: 5
bibliography: References.bib
csl: IJSEM.csl
---
```{r Packages, echo = FALSE, message = FALSE, warning=FALSE}
library("knitr")
library("tidyverse")
library("ggplot2")
library("ggpubr")
library("RColorBrewer")
library("extrafont")
theme_set(theme_classic())
```

@@

The closest relatives to both of the *Mixta* species were extracted from RM and BAM distance matrices. When the RM was used. *P. septica* was the closest relative to 66.3 and 66.5% of the *M. calida* and *M. gaviniae* sequences, respectively (Figure 1). These percentages increase by 12.6 and 14.4% for the *M. calida* and *M. gaviniae* sequences, respectively, when the BAM was used (Figure 1). Instead of *P. septica* when the RM is used, more genes were more closely related to *P. agglomerans*, *E. tasmaniensis*, *E. amylovora*, *T. saanichensis*, and *E. cloacae* (Figure 1). Additionally, there was more disparity between *M. calida* and *M. gaviniae* when the BAM was used (Figure 1).

@@

## Closest Relative from Genetic Distance
Genetic distance gives a numerical estimate of evolutionary change between species, and the taxa with which a species of interest has the shortest distance is regarded as the closest relative. Since the phylogenetic trees were calculated using the RM, the distance matrices exported from the trees were kept for future analyses. *P. septica* was consistently the closest relative for most of the *Mixta* sequences (Figure 2). However, the number of genes most closely related *P. septica* decreased from 66.3 and 66.5% (*M. calida* and *M. gaviniae*, respectively) with the NTS by 23.7% with the amino acid sequences (AAS; Figure 2). Instead, the *Mixta* AAS were most closely related to the other taxa within the dataset except for *P. syringae* (Figure 2). Interestingly, the number of genes most closely related to *E. cloacae* increased between the NTS and the AAS from 2.9% by 3.1% (Figure 2). Additionally, 55.1 and 55.2% of the *M. calida* and *M. gaviniae* sequences, respectively, had the same closest relatives between the NTS and the AAS analyses.

@@

## Spatial Patterns in the Closest Relatives
Circular plots were generated using the *Mixta* gene IDs to simulate their relative positions on the *Mixta* chromosomes and uncover positional and spatial patterns. The bars’ lengths and colours correspond to the species that was the closest relative to the *Mixta* gene at that position. The results between the two *Mixta* species for the closest relatives were the same except for one and two genes for the NTS and the AAS, respectively; therefore, only the results for *M. calida* are shown.

Many of the genes most closely related to any of the other taxa in this study were evenly positioned around the *Mixta* chromosome in the NTS (Figure 3A). The exception is for the *Tatumella* species (pinks) and *E. amylovora* (dark green; Figure 3A). The former and most of the genes closely related to the latter were located on the right side of the chromosome (Figure 3A). Unexpectedly, genes closely related to *E. cloacae* (orange) were distributed throughout the *Mixta* genome.

```{r Figure3, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 3. The closest relatives to *M. calida* genes for (**A**) nucleotide and (**B**) amino acid sequences around the *M. calida* chromosome. The length and colour of the bars represent the different species to which the gene is most closely related according to genetic distance. The *M. calida* genome has 4092 genes.", fig.width=6.5, fig.height=3.9}
FR_nucl_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure3_Circular_NT.csv", stringsAsFactors = FALSE) %>%
  mutate(Closest_Relative = factor(Closest_Relative, 
                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
                                              "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE))

FR_aa_calida <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure4_Circular_AA.csv", stringsAsFactors = FALSE) %>%
  mutate(Closest_Relative = factor(Closest_Relative, 
                                   levels = c("P_septica", "P_agglomerans", "E_tasmaniensis", "E_amylovora", "T_ptyseos", 
                                              "T_saanichensis", "E_cloacae", "P_syringae"), ordered = TRUE))

cal_NT <- ggplot(FR_nucl_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 20) +
  coord_polar() +
  scale_fill_brewer(palette = "Paired", 
                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
                               "P. syringae")) +
  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
  labs(fill = "Closest Relative") +
  theme_bw() +
  theme(legend.text = element_text(face = "italic"),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        text = element_text(size = 12,  family = "Times New Roman"))

cal_AA <- ggplot(FR_aa_calida, aes(x = ID, y = Relative_Number, fill = Closest_Relative)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 20) +
  coord_polar() +
  scale_fill_brewer(palette = "Paired", 
                    labels = c("P. septica", "P. agglomerans", "E. tasmaniensis", "E. amylovora", "T. ptyseos", "T. saanichensis", "E. cloacae",
                               "P. syringae")) +
  scale_y_continuous(limits = c(0, 9), breaks = c(0, 2, 4, 6, 8)) +
  labs(fill = "Closest Relative") +
  theme_bw() +
  theme(legend.text = element_text(face = "italic"),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        text = element_text(size = 12,  family = "Times New Roman"))

ggarrange(cal_NT, cal_AA, 
          labels = c("A", "B"),
          font.label = list(family = "Times New Roman"),
          label.x = 0.04,
          label.y = 0.98,
          common.legend = TRUE, 
          legend = "bottom")
```

In the AAS analysis, the genes closely related to the *Tatumella* species were not exclusive to the right half (Figure 3B). Additionally, the genes closely related to the *Erwinia* species and *E. cloacae* were more evenly distributed across the genome (Figure 3B). Similar to the NTS analysis, no gene clusters related to a single taxon were evident (Figure 3B).

@@

<!-- ```{r Figure1, echo = FALSE, message = FALSE, warning=FALSE, fig.cap="Figure 1. The percent of genes (n = 799) requiring each model when using the recommended model for phylogenetic tree analysis (green) and when using the best available model for genetic distance analysis (orange). Models are for nucleotide sequences.", fig.width=6.5} -->
<!-- models <- read.csv("C:/Users/Kim/OneDrive/2020_3Fall/Biology_396/10_Tables_and_Figures/Figure1_BestvAvailableModels.csv", stringsAsFactors = FALSE) -->

<!-- ggplot(data = models, aes(x = Model, y = Percent, fill = Bst_Avlb)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
<!--   scale_fill_brewer(palette = "Dark2", -->
<!--                     labels = c("Phylogenetic", "Genetic Distance")) + -->
<!--   labs(y = "Percent of Genes \n Requiring Each Model", -->
<!--        fill = "Analysis") + -->
<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1),                                # Rotates x axis labels -->
<!--         legend.position = c(0.3, 0.95),                                                   # Moves legend inside the plot and in the top-left corner -->
<!--         legend.justification = c("right", "top"),                                         # Top right corner is at above coordinates -->
<!--         legend.text.align = 0,                                                            # Aligns legend text to the left -->
<!--         legend.margin = margin(6, 6, 6, 6),                                               # Margins around legend -->
<!--         text = element_text(size = 12,  family = "Times New Roman")) -->
<!-- ``` -->
